Weighted Model Counting Without Parameter Variables

### 2021-02-14 doc/paper2/paper.tex
Some writing and editing

The most well-studied version of WMC assigns weights to models based on
weights on literals, i.e., the weight of a model is the product of the weights
of all literals in it. This simplification is motivated by the fact that the
number of models scales exponentially with the number of atoms, so listing the
weight of every model is intractable. However, this also severely restricts what
probability distributions can be represented. A common way to overcome this
limitation is by adding more literals. While we show that this is always
possible, we also demonstrate that it can be significantly more efficient to
encode weights in a more flexible format instead.

@@

This new measure-theoretic perspective allows us to not only encode any
discrete probability distribution but also improve inference speed. In
\cref{sec:bns}, we demonstrate this by developing a new WMC encoding for
Bayesian networks that uses \emph{conditional weights} on literals (in the
spirit of conditional probabilities) that have literal-based WMC as a special
case. We prove the correctness of the encoding and show how a state-of-the-art
WMC solver \textsf{ADDMC} \citep{DBLP:conf/aaai/DudekPV20} can be adapted to the
new format. \textsf{ADDMC} is a recently-proposed algorithm for WMC based on
manipulating functions on Boolean algebras using an efficient representation for
such functions known as algebraic decision diagrams (ADDs)
\citep{DBLP:journals/fmsd/BaharFGHMPS97}. Finally, in \cref{sec:experiments}, we
show how the new encoding results in faster inference on the vast majority of
benchmark instances, often by one or two orders of magnitude. We explain the
performance benefits by showing how our encoding has asymptotically fewer
variables and ADDs.

@@

\citet{DBLP:conf/aaai/SangBK05} claim that
\texttt{sbk05} is a smaller encoding than \texttt{d02} with respect to both the
number of clauses and the number of variables but provide no experimental
comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with
\texttt{d02} by measuring the time it takes to compile either encoding into an
arithmetic circuit. They show that
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\citep{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \citep{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with \textsf{Cachet} \citep{DBLP:conf/sat/SangBBKP04} with
the compile times of \texttt{cd05} and \texttt{cd06}, but only on five (types
of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Finally,
\citet{DBLP:conf/ecai/BartKLM16} introduce \texttt{bklm16} and show that it has
both fewer variables and fewer clauses than \texttt{cd06}. Their experiments
show \texttt{bklm16} to be superior to \texttt{cd06} with respect to both
compilation time and encoding size when both are compiled using
\textsf{c2d}\footnote{\url{http://reasoning.cs.ucla.edu/c2d/}}
\citep{DBLP:conf/ecai/Darwiche04} but inferior to \texttt{cd06} when
\texttt{cd06} is compiled using
\textsf{Ace}\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} (which still uses
\textsf{c2d} but considers the structure of the Bayesian network along with its
encoding). Our experiments in \cref{sec:experiments} confirm some of the
findings outlined in this section while also observing that the (relative and
absolute) performance of each encoding depends on the WMC algorithm in use, and
smaller encodings are not necessarily faster.

@@

We attempt to explain what makes \textsf{ADDMC} run hundreds of times faster
with \texttt{cw} than with any other encoding by considering asymptotic upper
bounds on the numbers of variables and ADDs based on the size and structure of
the Bayesian network. Let $n = |\mathcal{V}|$ be the number of vertices in the
Bayesian network, $d = \max_{X \in \mathcal{V}} |\mathrm{pa}(X)|$ the maximum
in-degree (i.e., the number of parents), and $v = \max_{X \in \mathcal{V}} |\im
X|$ the maximum number of values per variable. \Cref{tbl:asymptotes} shows how
\texttt{cw} has fewer variables and fewer ADDs than any other encoding. We
conjecture that it is primarily the reduced number of variables that makes the
\textsf{ADDMC} variable ordering heuristics much more effective. Note that these
are upper bounds and most encodings (including \texttt{cw}) can be smaller in
certain situations (e.g., with binary random variables or when a CPT has
repeating probabilities). We equate clauses and ADDs (more specifically, factors
of the function $\phi$ from \cref{alg:encoding}) here because \textsf{ADDMC}
interprets each clause of any WMC encoding as a multiplicative factor of the ADD
that represents the entire WMC instance \citep{DBLP:conf/aaai/DudekPV20}. For
literal-weight encodings, each weight is also a factor, but that does not affect
our asymptotic bounds.

>>>>

The most well-known version of WMC assigns weights to models based on
weights on literals, i.e., the weight of a model is the product of the weights
of all literals in it. This simplification is motivated by the fact that the
number of models scales exponentially with the number of atoms, so listing the
weight of every model is intractable. However, this also severely restricts what
probability distributions can be represented. A common way to overcome this
limitation is by adding more literals. While we show that this is always
possible, we demonstrate that it can be significantly more efficient to
encode weights in a more flexible format instead.

@@

This new perspective allows us to not only encode any discrete probability
distribution but also improve inference speed. In \cref{sec:bns} we demonstrate
this by developing a new WMC encoding for Bayesian networks that uses
\emph{conditional weights} on literals (in the spirit of conditional
probabilities) that have literal-based WMC as a special case. We prove the
correctness of the encoding and show how a state-of-the-art WMC solver
\textsf{ADDMC} \citep{DBLP:conf/aaai/DudekPV20} can be adapted to the new
format. \textsf{ADDMC} is a recently-proposed algorithm for WMC based on
manipulating functions on Boolean algebras using an efficient representation for
such functions known as algebraic decision diagrams (ADDs)
\citep{DBLP:journals/fmsd/BaharFGHMPS97}. \textsf{ADDMC} was already shown to be
capable of solving instances other solvers fail at and being the fastest solver
on the largest number of instances \citep{DBLP:conf/aaai/DudekPV20}. Our
experiments in \cref{sec:experiments} focus on further improving the performance
of \textsf{ADDMC} on instances that originate from Bayesian networks. We show
how our new encoding improves inference on the vast majority of benchmark
instances, often by one or two orders of magnitude. We explain the performance
benefits by showing how our encoding has asymptotically fewer variables and
ADDs.

@@

\citet{DBLP:conf/aaai/SangBK05} claim that
\texttt{sbk05} is a smaller encoding than \texttt{d02} with respect to both the
number of clauses and the number of variables but provide no experimental
comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with
\texttt{d02} by measuring the time it takes to compile either encoding into an
arithmetic circuit. They show that
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\citep{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \citep{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with \textsf{Cachet} \citep{DBLP:conf/sat/SangBBKP04} with
the compile times of \texttt{cd05} and \texttt{cd06}, but only on five (types
of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed.  The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Finally, \citet{DBLP:conf/ecai/BartKLM16} introduce
\texttt{bklm16} and show that it has both fewer variables and fewer clauses than
\texttt{cd06}. Their experiments show \texttt{bklm16} to be superior to
\texttt{cd06} with respect to both compilation time and encoding size when both
are compiled using
\textsf{c2d}\footnote{\url{http://reasoning.cs.ucla.edu/c2d/}}
\citep{DBLP:conf/ecai/Darwiche04} but inferior to \texttt{cd06} when
\texttt{cd06} is compiled using
\textsf{Ace}\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} (which still uses
\textsf{c2d} but considers the structure of the Bayesian network along with its
encoding). Our experiments in \cref{sec:experiments} confirm some of the
findings outlined in this section while also showing that the performance of
each encoding depends on the WMC algorithm in use, and smaller encodings are not
necessarily faster.

@@

We can explain what makes \textsf{ADDMC} run significantly faster with
\texttt{cw} than with any other encoding by considering asymptotic upper bounds
on the numbers of variables and ADDs based on the size and structure of the
Bayesian network. Let $n = |\mathcal{V}|$ be the number of vertices in the
Bayesian network, $d = \max_{X \in \mathcal{V}} |\mathrm{pa}(X)|$ the maximum
in-degree (i.e., the number of parents), and $v = \max_{X \in \mathcal{V}} |\im
X|$ the maximum number of values per variable. \Cref{tbl:asymptotes} shows how
\texttt{cw} has fewer variables and fewer ADDs than any other encoding. We
conjecture that it is primarily the reduced number of variables that makes the
\textsf{ADDMC} variable ordering heuristics much more effective. Note that these
are upper bounds and most encodings (including \texttt{cw}) can be smaller in
certain situations (e.g., with binary random variables or when a CPT has
repeating probabilities). We equate clauses and ADDs (more specifically, factors
of the function $\phi$ from \cref{alg:encoding}) here because \textsf{ADDMC}
interprets each clause of any WMC encoding as a multiplicative factor of the ADD
that represents the entire WMC instance \citep{DBLP:conf/aaai/DudekPV20}. For
literal-weight encodings, each weight is also a factor, but that does not affect
our asymptotic bounds.

### 2021-02-14 doc/paper3/paper.tex
Some writing and editing

% TODO: make introduce an acronym: project-join tree = PJT
\section{Main Results}

\paragraph{Notation.} For any propositional formula $\phi$ and $p, q \in
\mathbb{R}_{\ge 0}$,
\[
  [\phi]^p_q \coloneqq
  \begin{cases}
    p & \text{if } \phi \\
    q & \text{otherwise.}
  \end{cases}
\]

\todo[inline]{Formal definition of a previous WMC instance (CNF, literal weight
  function) and the new definition (set of $2^{X} \to \mathbb{R}_{\ge 0}$
  pseudo-Boolean functions and a constant). Note that the constant idea is
  borrowed from \texttt{bklm16}.}

\paragraph{References.}
\begin{itemize}
\item `Apply' is quadratic \cite{DBLP:journals/tc/Bryant86}.
\item Related work without publicly available implementations:
  \begin{itemize}
  \item direct compilation to SDDs \cite{DBLP:conf/ecsqaru/ChoiKD13}
  \item direct compilation to PSDDs, also eliminating parameter variables (a thesis)
  \item maybe two more papers
  \end{itemize}
\end{itemize}
% TODO: do a literature search focused around these papers

\paragraph{Notes.}
\begin{itemize}
\item Let $X_P$ be the set of parameter variable and $X_I$ be the set of
  indicator variables.
\item Parameter variables are either taken from the LMAP file (for encodings
  produced by Ace) or assumed to be the variables that have both weights equal
  to 1.
\item If a parameter variable in a clause is `negated', we can ignore the
  clause. We assume that there are no clauses with more than one instance of
  parameter variables.
\item The second \textbf{foreach} loop can be performed in constant time by
  representing $\phi'$ as a list and assuming that the two 'clauses' are
  adjacent in that list (and incorporating it into the first loop).
\item The $d$ map is constructed in $\mathcal{O}(|X_P|\log|X_P|)$ time (we want
  to use a data structure based on binary search trees rather than hashing).
\item \texttt{rename} can be implemented in $\mathcal{O}(\log |X_P|)$ time.
\item Apparently, the DPMC paper already shows that taking the first offered
  decomposition tree is best.
\item This may look like preprocessing, but all the transformations are local
  and thus can be incorporated into an encoding algorithm with no slowdown. In
  fact, if anything, the resulting algorithm would be slightly faster, as it
  would have less data to output.
\item It is already well-known that WMC is FPT.
\end{itemize}

@@

% NOTE: By DPMC, we always mean DMC+lg.
% TODO: maybe use texttt for DPMC?

Summary:
\begin{itemize}
\item We establish DPMC inference as fixed-parameter tractable.
\item We show that, in the case of cw, the parameter is the same as in the BN lower bound,
\item and, in the case of d02, the DPMC parameter is exponentially higher.
\item We experimentally show that cw+DPMC is best on low-to-moderate treewidth
  instances, and cd06+c2d overtakes cw+DPMC on higher treewidth instances.
\end{itemize}

Define:
\begin{itemize}
\item treewidth of a graph,
\item Boolean formula in CNF (perhaps this is too trivial to define),
\item primal graph of a CNF formula (a.k.a.
  Gaifman/(variable) interaction/connectivity/clique/representing graph),
\item Bayesian network (I already have a definition of these last two),
\item moralisation of a Bayesian network (or of any DAG),
\end{itemize}

@@

\begin{theorem}[\cite{DBLP:conf/cp/DudekPV20}, paraphrased, `ADD width is equal
  to the treewidth of the primal graph']
  Given a CNF formula $\phi$ with a tree decomposition of its primal graph of
  width $w$, Algorithm~2 (from \cite{DBLP:conf/cp/DudekPV20}) returns a
  PJT of $\phi$ of width at most $w+1$.
\end{theorem}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[circle,draw] {$\exists$}
    child {node (a) {$\vdots$}
      child {node[circle,draw] {$c_1$}}
    }
    child {node (b) {$\vdots$}
      child {node[circle,draw] {$\ast$}
        child {node {$\vdots$}
          child {node[circle,draw] {$c_i$}}
        }
      }
    }
    child {node (c) {$\vdots$}
      child {node[circle,draw] {$c_l$}}
    }
    ;
    \path (a) -- (b) node[midway] {$\cdots$};
    \path (b) -- (c) node[midway] {$\cdots$};
  \end{tikzpicture}
  \caption{An example PJT with clauses $c_1, \dots, c_i, \dots, c_l$ that all
    contain the same variable, its projection node $\exists$, and a node under
    consideration $\ast$.}
\end{figure}

\begin{lemma}
  The number of variables that can appear in the ADDs within a PJT node is $\le
  k+1$.
\end{lemma}
\begin{proof}
  All such variables are introduced in a `clause' below and projected either in
  this node or in one of its ancestors. Each such variable would be in the bag
  of each clause that introduced it. By the tree decomposition properties, it
  will also be in the bag of both the projection node and the current node
  because they are on the path between the clause nodes (see the figure).
\end{proof}

>>>>

\section{Introduction}

\paragraph{Notation.} For any propositional formula $\phi$ and $p, q \in
\mathbb{R}_{\ge 0}$, let $[\phi]^p_q\colon 2^X \to \mathbb{R}$ be a
pseudo-Boolean function defined as
\[
  [\phi]^p_q \coloneqq
  \begin{cases}
    p & \text{if } \phi \\
    q & \text{otherwise.}
  \end{cases}
\]

\paragraph{References.}
\begin{itemize}
\item Related work without publicly available implementations:
  \begin{itemize}
  \item direct compilation to SDDs \cite{DBLP:conf/ecsqaru/ChoiKD13}
  \item direct compilation to PSDDs, also eliminating parameter variables (a thesis)
  \item maybe two more papers
  \end{itemize}
\end{itemize}
% TODO: do a literature search focused around these papers
% TODO : use mathcal on V, L, E (and introduce them)
% TODO: maybe use texttt for DPMC?

\paragraph{Notes.}
\begin{itemize}
\item Apparently, the DPMC paper already shows that taking the first offered
  decomposition tree is best.
\item It is already well-known that WMC is FPT.
\end{itemize}

\section{Parameter Variable Elimination}

\paragraph{Notes.}
\begin{itemize}
\item Let $X_P$ be the set of parameter variable and $X_I$ be the set of
  indicator variables.
\item Parameter variables are either taken from the LMAP file (for encodings
  produced by Ace) or assumed to be the variables that have both weights equal
  to 1.
\item If a parameter variable in a clause is `negated', we can ignore the
  clause. We assume that there are no clauses with more than one instance of
  parameter variables.
\item The second \textbf{foreach} loop can be performed in constant time by
  representing $\phi'$ as a list and assuming that the two 'clauses' are
  adjacent in that list (and incorporating it into the first loop).
\item The $d$ map is constructed in $\mathcal{O}(|X_P|\log|X_P|)$ time (we want
  to use a data structure based on binary search trees rather than hashing).
\item \texttt{rename} can be implemented in $\mathcal{O}(\log |X_P|)$ time.
\item This may look like preprocessing, but all the transformations are local
  and thus can be incorporated into an encoding algorithm with no slowdown. In
  fact, if anything, the resulting algorithm would be slightly faster, as it
  would have less data to output.
\end{itemize}

@@

\paragraph{Notes.}
\begin{itemize}
\item Summary of results
  \begin{itemize}
  \item We establish DPMC inference as fixed-parameter tractable.
  \item We experimentally show that cw+DPMC is best on low-to-moderate treewidth
    instances, and cd06+c2d overtakes cw+DPMC on higher treewidth instances.
  \end{itemize}
\item By DPMC, we always mean DMC+lg.
\end{itemize}

\paragraph{TODO: Define:}
\begin{itemize}
\item Formal definition of a previous WMC instance (CNF, literal weight
  function) and the new definition (set of $2^{X} \to \mathbb{R}_{\ge 0}$
  pseudo-Boolean functions and a constant). Note that the constant idea is
  borrowed from \texttt{bklm16}.
\item Boolean formula in CNF (perhaps this is too trivial to define),
\item primal graph of a CNF formula (a.k.a.
  Gaifman/(variable) interaction/connectivity/clique/representing graph),
\item Bayesian network (I already have a definition of these last two),
\item moralisation of a Bayesian network (or of any DAG),
\end{itemize}

@@

\begin{theorem}[\cite{DBLP:conf/cp/DudekPV20}, rephrased, `ADD width is equal
  to the treewidth of the primal graph']
  Given a CNF formula $\phi$ with a tree decomposition of its primal graph of
  width $w$, Algorithm~2 (from \cite{DBLP:conf/cp/DudekPV20}) returns a
  PJT of $\phi$ of width at most $w+1$.
\end{theorem}

\begin{definition}[\cite{DBLP:journals/jct/RobertsonS84}, rephrased]
  A \emph{tree decomposition} of a graph $G$ is a pair $(T, \chi)$, where $T$ is
  a tree and $\chi\colon V(T) \to 2^{V(G)}$ is a labelling function, with the
  following properties:
  \begin{itemize}
  \item $\bigcup_{t \in V(T)} \chi(t) = V(G)$;
  \item for every edge $e \in E(G)$, there exists $t \in V(T)$ such that $e$ has
    both endpoints in $\chi(t)$;
  \item for all $t, t', t'' \in V(T)$, if $t'$ is on the path between $t$ and
    $t''$, then $\chi(t) \cap \chi(t'') \subseteq \chi(t')$.
  \end{itemize}
  The \emph{width} of tree decomposition $(T, \chi)$ is $\max_{t \in V(T)}
  |\chi(t)| - 1)$. The \emph{treewidth} of graph $G$ is the smallest $w$ such
  that $G$ has a tree decomposition of width $w$.
\end{definition}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[circle,draw] {$\exists$}
    child {node (a) {$\vdots$}
      child {node[circle,draw] {$c_1$}}
    }
    child {node (b) {$\vdots$}
      child {node[circle,draw] {$\ast$}
        child {node {$\vdots$}
          child {node[circle,draw] {$c_i$}}
        }
      }
    }
    child {node (c) {$\vdots$}
      child {node[circle,draw] {$c_l$}}
    }
    ;
    \path (a) -- (b) node[midway] {$\cdots$};
    \path (b) -- (c) node[midway] {$\cdots$};
  \end{tikzpicture}
  \caption{An example PJT with clauses $c_1, \dots, c_i, \dots, c_l$ that all
    contain the same variable, its projection node $\exists$, and a node under
    consideration $\ast$.}
  \label{fig:example_pjt}
\end{figure}

\begin{definition}
  Let $f, g\colon 2^X \to \mathbb{R}$ be pseudo-Boolean functions. Operations
  such as addition and multiplication are defined pointwise as
  \[
    (f+g)(Y) \coloneqq f(Y)+g(Y),
  \]
  and
  \[
    (fg)(Y) \coloneqq f(Y)g(Y)
  \]
  for all $Y \subseteq X$.
\end{definition}
% TODO: do I need to formally consider extending a pseudo-Boolean function to a
% bigger domain?

\begin{definition}
  Let $f\colon 2^X \to \mathbb{R}$ be a pseudo-Boolean function, and $x \in X$.
  Then $f|_{x=0}, f|_{x=1}\colon 2^X \to \mathbb{R}$ are \emph{restrictions} of
  $f$ defined as
  \[
    f|_{x=0}(Y) \coloneqq f(Y \setminus \{x\}),
  \]
  and
  \[
    f|_{x=1}(Y) \coloneqq f(Y \cup \{x\})
  \]
  for all $Y \subseteq X$.
\end{definition}

\begin{definition} \label{def:projection}
  Let $X$ be a set. For any $x \in X$, \emph{projection} $\exists_x$ is an
  endomorphism $\exists_x\colon \mathbb{R}^{2^X} \to \mathbb{R}^{2^X}$ defined
  as
  \[
    \exists_xf = f|_{x=1} + f|_{x=0}
  \]
  for any $f\colon 2^X \to \mathbb{R}$.
\end{definition}
% TODO: gamma will have to point to ADDs instead (and not be a bijection)

\begin{lemma}
  Let $f,g\colon 2^X \to \mathbb{R}$ be pseudo-Boolean functions represented by
  ADDs with $n$ and $m$ nodes, respectively, and $x \in X$. Then $f+g$ and $fg$
  can be computed in $\mathcal{O}(mn)$ time, and $\exists_xf$ can be computed in
  $\mathcal{O}(n^2)$ time.
\end{lemma}
\begin{proof}
  Addition and multiplication are implemented by \texttt{apply} algorithm which
  takes $\mathcal{O}(mn)$ time \cite{DBLP:journals/tc/Bryant86}. Projection
  consists of two restrictions and an addition by \cref{def:projection}. The
  computational complexity of restriction is dominated by the reduction
  operation that transforms a decision diagram into a minimal canonical form
  \cite{DBLP:journals/tc/Bryant86}. While the original reduction algorithm had a
  $\mathcal{O}(n \log n)$ complexity, caching can reduce it to $\mathcal{O}(n)$
  \cite{somenzi1998cudd}. Either way, the complexity of projection is still
  $\mathcal{O}(n^2)$.
\end{proof}

\begin{definition}
  Let $(T, r, \gamma, \pi)$ be a PJT, and $X$ be the set of variables. The
  functionality of DPMC execution (i.e., Algorithm~1 in the ADDMC paper
  \cite{DBLP:conf/aaai/DudekPV20}) can be represented by $\delta(r)$, where
  $\delta\colon \mathcal{V}(T) \to \mathbb{R}^{2^X}$ is a recursive function
  defined as
  \[
    \delta(t) =
    \begin{cases}
      \gamma(t) & \text{if } t \in \mathcal{L}(T) \\
      \exists_{\pi(t)} \prod_{u \in \mathcal{C}(t)} \delta(u) & \text{otherwise.}
    \end{cases}
  \]
  The range of $\delta(r)$ then contains a single real number, i.e., the answer.
\end{definition}

\begin{lemma}
  The number of variables that can appear in the ADDs within a PJT node is $\le
  k+1$.
\end{lemma}
\begin{proof}
  All such variables are introduced in a `clause' below and projected either in
  this node or in one of its ancestors. Each such variable would be in the bag
  of each clause that introduced it. By the tree decomposition properties, it
  will also be in the bag of both the projection node and the current node
  because they are on the path between the clause nodes (see
  \cref{fig:example_pjt}).
  \todo[inline]{How do I formalise this argument?}
\end{proof}
